\LARGE A double-sampling extension of the German National Forest Inventory for design-based small area regression estimation on forest district levels \\ \vspace{2ex} \large Large scale application to the federal state of Rhineland-Palatinate \vspace{-1ex}

### 2018-02-27 Draft_Form/tex/Results.tex
update 12_02_2018
update 12_02_2018
update 27_02_2018
update 27_02_2018 (Alex review)

An application of the SRS, \psmall{} and \extpsynth{} estimator was not feasible for 17 of all 405 FR-units due to an insufficient terrestrial sample size of $n_{2,G} < 2$. We further restricted the calculation of the \psmall{} and \extpsynth{} estimator to small area units with a minimum terrestrial sample size of $n_{2,G} \geq 4$ to avoid unstable estimates. This affected 65 additional FR units and limited unbiased two-phase estimations to 321 (79\%) of the 405 FR units. It should be noted that also the \psynth{} estimator could not be applied for 2 FR-units since $n_{1,G} < 2$. Due to substantial larger sample sizes, all estimators could however be applied to all 45 FA units. The average value and the range of the mean timber volume estimates over the evaluated FA and FR units turned out to be very similar between all estimators (table \ref{tab:estres}). An additional pairwise comparison of the 95\% confidence intervals revealed that the four estimators did in fact not produce statistically different point estimates for all FA and FR units. This confirmed that the differences between the estimators are solely found in the precision which they provide for the point estimates.

@@

\begin{table}[H]
	\begin{center}
		\caption{Descriptive summary of point estimates and estimation errors on the two forest district levels. $N_u$: number of evaluated small area units.}
		\vspace{0.2cm}
		\label{tab:estres}
		{\small %
			\begin{tabular}{c|l c|c c c|c c c} %8cols
				\hlineB{1}
				\multirow{2}{*}{District level} & \multicolumn{2}{c|}{\multirow{2}{*}{Estimator}} & \multicolumn{3}{c|}{Point estimates} & \multicolumn{3}{c}{Errors [\%]} \\
				\cline{4-9} & & & mean & min & max & mean & min & max \\
				\hline \hline
				\multirow{4}{*}{FA} & SRS       & ($N_u$=45)  & 300.16 & 215.91 & 392.84 &  6.69 & 3.87 & 13.21 \\
				                    & PSMALL    & ($N_u$=45)  & 307.29 & 209.26 & 417.10 &  5.16 & 3.46 & 14.33 \\
				                    & EXTPSYNTH & ($N_u$=45)  & 307.27 & 209.01 & 415.02 &  4.78 & 3.25 & 13.88 \\
				                    & PSYNTH    & ($N_u$=45)  & 306.90 & 223.51 & 409.92 &  2.34 & 1.54 &  3.95 \\
				\hlineB{2}          
				\multirow{4}{*}{FR} & SRS       & ($N_u$=388) & 301.83 &  99.89 & 612.13 & 18.32 & 0.34 & 104.97 \\
				                    & PSMALL    & ($N_u$=321) & 308.15 & 159.64 & 568.67 & 12.24 & 3.48 &  44.94 \\
				                    & EXTPSYNTH & ($N_u$=321) & 308.38 & 154.07 & 544.34 & 11.34 & 3.60 &  40.91 \\
				                    & PSYNTH    & ($N_u$=403) & 307.82 & 166.01 & 444.29 & 4.65  & 2.56 &  62.51 \\
				\hline \hline
			\end{tabular}
		}%
	\end{center}
\end{table}

% ----------------------------------------------------------------------- %
% ----------------------------------------------------------------------- %
\subsection{Estimation errors}
\label{sec:esterr}

On both small area levels, the design-unbiased estimators \psmall{} and \extpsynth{} led to a substantial reduction in the estimation errors compared to the SRS estimator (fig. \ref{fig:disterrors}). On the FA level, the SRS estimator yielded an estimation error of 6.7\% on average compared to 5.2\% and 4.8\% under the \extpsynth{} and \psmall{} estimator (table \ref{tab:estres}). The cumulative error distribution (fig. \ref{fig:disterrors}, left) reveals that under the SRS estimator, errors $\leq$ 5\% were achieved for 17\% of the FA units (8 of 45). This proportion could be increased to 62\% (28 FA units) and 73\% (33 FA units) by application of the \psmall{} and \extpsynth{} estimator. 95\% of all estimates exhibited errors $\leq$ 9.5\% under the SRS estimator and $\leq$ 6.6\% when using \psmall{} or \extpsynth{}. Estimation errors higher than 10\% only appeared twice for each of the three estimators.\par
The error reduction by the \psmall{} and \extpsynth{} estimator was even more pronounced on the FR level (fig. \ref{fig:disterrors}, right), although the overall error niveau was substantially higher than on FA level. The average error under the SRS estimator was 18.3\%, while it was 11.3\% and 12.2\% under the \psmall{} and \extpsynth{} estimator (table \ref{tab:estres}). Errors smaller than 10\% were achieved for 15\% of the FR units by the SRS estimator, and for 46\% by the \psmall{} and \psynth{} estimator. 95\% of the 321 FR units where \psmall{} and \extpsynth{} could be applied exhibited errors $\leq$ 20\%. In comparion, the SRS estimates resulted in errors $\leq$ 36.6\% for 95\% of the 388 FR units.

\begin{figure}[H]
	\centering
	\resizebox{1\hsize}{!}{\includegraphics{fig/error_distr_foa_fu.png}}
	\caption{Cumulative distribution of estimation errors under the simple random sampling (SRS), the pseudo small (\psmall{}), the extended pseudo synthetic (\extpsynth{}) and the pseudo synthetic (\psynth{}) estimator. \textit{Left}: Results for the 45 FA units. \textit{Right}: Results for the 388 (SRS), 321 (\psmall{} / \extpsynth{}) and 403 (\psynth{}) FR units.}
	\label{fig:disterrors}
\end{figure}

@@

Figure \ref{fig:disterrors} reveals that the error distribution of \psmall{} and \extpsynth{} are very similar, with \psmall{} showing marginally higher estimation errors. In order to investigate the differences between \psmall{} and \extpsynth{}, we compared the g-weight variances of both estimators for all 321 FR units (fig. \ref{fig:compvar}, left). As obvious, \psmall{} yielded slightly larger variances for the vast majority of the estimates. As addressed in section \ref{sec:extpsynth}, one possible explanation for such differences was the effect of one or more cluster not entirely being included in a small area unit, as this would constitute a violation of the \extpsynth{} estimator. This violation was actually observed in 155 of the 321 FR units (48\%). However, the affected FR units (depicted in red diamonds, fig. \ref{fig:compvar}) did not show a significant divergence from the \psmall{} variances with respect to the remaining unaffected FR units. The variance differences between the two estimators were thus due to the mathematical formulations of the \psmall{} and \extpsynth{} estimator, which are asymptotically equivalent only under large terrestrial sample sizes $n_{2,G}$ within the small area \citep[pp.17--18]{mandallaz2016}. An additional comparison of the absolute differences in the g-weight variance (fig. \ref{fig:compvar}, right) revealed that large divergences did in fact particularly occur for small area units with small terrestrial sample sizes ($n_{2,G} \leq 5$). The differences decreased with increasing sample size and thus confirmed the asymptotic relationship between the two estimators. However, a comparison of the confidence intervals of \psmall{} and \extpsynth{} revealed that the variance differences did not lead to statistically significant point estimates.\par

@@

% ----------------------------------------------------------------------- %
% ----------------------------------------------------------------------- %
\subsection{Reduction of SRS variance by \psmall{} and \extpsynth{}}
\label{sec:gain_eval}

A direct comparison of the realized variances within the small area units revealed that the application of the design-unbiased estimators (\psmall{} and \extpsynth{}) led to a reduction of the respective SRS variance in all FA units. In 75\% of the FA units, the \extpsynth{} estimator was able to reduce the SRS variance by up to 54.1\% (fig. \ref{fig:gain}). The reduction in variance was also expressed in the relative efficiency values, which were 2.02 on average and ranged between 1.18 and 4.13 on the FA level. On FR-level, the reduction in variance and the relative efficiencies reached even higher values (table \ref{tab:gain} and fig. \ref{fig:gain}). The \psmall{} estimator again yielded slightly lower variance reductions and relative efficiencies due to the generally smaller variances of the \extpsynth{} estimator (section \ref{sec:comp}).

\begin{figure}[H]
	\centering
	\resizebox{0.7\hsize}{!}{\includegraphics{fig/compare_onephase_extpsynth_psmall.png}}
	\caption{Cumulative distribution of variance reduction by the PSMALL and EXTPSYNTH compared to the SRS estimator for the  45 FA and 321 FR units.}
	\label{fig:gain}
\end{figure}

\begin{table}[H]
	\begin{center}
		\caption{Descriptive summary of SRS variance reduction and relative efficiencies on the two forest district levels. $N_u$: number of evaluated small area units.}
		\vspace{0.2cm}
		\label{tab:gain}
		{\small %
			\begin{tabular}{c|l c|c c c|c c c} %8cols
				\hlineB{1}
				\multirow{2}{*}{District level} & \multicolumn{2}{c|}{\multirow{2}{*}{Estimator}} & \multicolumn{3}{c|}{Reduction of SRS variance [\%]} & \multicolumn{3}{c}{relative efficiency} \\
				\cline{4-9} & & & mean & min & max & mean & min & max \\
				\hline \hline
				\multirow{2}{*}{FA} & PSMALL    & ($N_u$=45)  & 33.51 &  2.6 & 72.5 & 1.74 & 1.03 & 3.64 \\
				& EXTPSYNTH & ($N_u$=45)  & 43.30 & 15.7 & 75.8 & 2.03  & 1.18 & 4.13 \\
				\hlineB{2}          
				\multirow{2}{*}{FR} & PSMALL    & ($N_u$=321) & 12.48 & -1203.9 & 96.8 & 2.54 & 0.08 & 31.61  \\
				& EXTPSYNTH & ($N_u$=321) & 24.75 & -892.7  & 97.0 & 2.95 & 0.10 & 33.70 \\
				\hline \hline
			\end{tabular}
		}%
	\end{center}
\end{table}

However, cases also occurred on the FR level where one or both two-phase estimators produced larger variance values than under the SRS estimator. This particularly happened in 19\% (61) of the FR units under the \extpsynth{}, and in 24\% (76) of the FR units under the \psmall{} estimator. One possible reason for this was supposed to be a large residual variance due to a poor performance of the regression model within the small area unit. In order to investigate this hypothesis, we analyzed the three variance terms of the \psmall{} estimator (eq. \ref{eq:psmall_var}), i.e. the variance introduced by the uncertainty of the regression coefficients (term 1), the variance caused by estimating the auxiliary means (term 2), and the variance of the model residual (term 3).  The latter particularly expresses the model performance within the small area unit. Figure \ref{fig:fail} illustrates the percentage reduction or increase of the SRS variance when compared to the \psmall{} variance for all FR units in dependence on a) the impact of residual variance on the g-weight variance, and b) the terrestrial sample size $n_{2,G}$.\par

Obviously, the residual term generally constitutes the dominating part of the \psmall{} g-weight variance (around 84\% on average). However, a high proportion of the residual variance term seems not to be the driver for large \psmall{} variances, as apparent from Fig. \ref{fig:fail} (\textit{right}). The FR units where the \psmall{} estimator produced larger variances than the SRS estimator did not systematically differ from the majority of the FR units where \psmall{} performed better than SRS. However, FR units with exceptionally large variance increases compared to SRS particularly occurred under small terrestrial sample sizes of $n_{2,G} = 4$ (fig. \ref{fig:fail}, \textit{left}). These FR units exhibited a 272\% average increase of the SRS variance, compared to 62\% for the critical FR units with $n_{2,G} > 4$. In comparison, the achievable reduction in SRS variances compared to SRS were not considerably impacted by the terrestrial sample size (fig. \ref{fig:fail}, \textit{right}). The average reduction of these units was around 50\% under sampling sizes of both $n_{2,G} = 4$ and $n_{2,G} > 4$.

>>>>

An application of the SRS, \psmall{} and \extpsynth{} estimator was not feasible for 17 of all 405 FR-units due to an insufficient terrestrial sample size of $n_{2,G} < 2$. We further restricted the calculation of the \psmall{} and \extpsynth{} estimator to small area units with a minimum terrestrial sample size of $n_{2,G} \geq 4$ to avoid unstable estimates. This affected 65 additional FR units and limited unbiased two-phase estimations to 321 (79\%) of the 405 FR units. It should be noted that also the \psynth{} estimator could not be applied for 2 FR-units since $n_{1,G} < 2$. Due to substantially larger sample sizes, all estimators could however be applied to all 45 FA units. The average value and the range of the mean timber volume estimates over the evaluated FA and FR units turned out to be very similar between all estimators (Table \ref{tab:estres}). An additional pairwise comparison of the 95\% confidence intervals revealed that the four estimators did in fact not produce statistically different point estimates for all FA and FR units. This confirmed that the differences between the estimators are solely found in the precision which they provide for the point estimates.

@@

\begin{table}[H]
	\begin{center}
		\caption{Descriptive summary of point estimates and estimation errors on the two forest district levels. $N_u$: number of evaluated small area units.}
		\vspace{0.2cm}
		\label{tab:estres}
		{\small %
			\begin{tabular}{c|l c|c c c|c c c} %8cols
				\hlineB{1}
				\multirow{2}{*}{District level} & \multicolumn{2}{c|}{\multirow{2}{*}{Estimator}} & \multicolumn{3}{c|}{Point estimates} & \multicolumn{3}{c}{error$_{[\%]}$} \\
				\cline{4-9} & & & mean & min & max & mean & min & max \\
				\hline \hline
				\multirow{4}{*}{FA} & SRS       & ($N_u$=45)  & 300.16 & 215.91 & 392.84 &  6.69 & 3.87 & 13.21 \\
				                    & PSMALL    & ($N_u$=45)  & 307.29 & 209.26 & 417.10 &  5.16 & 3.46 & 14.33 \\
				                    & EXTPSYNTH & ($N_u$=45)  & 307.27 & 209.01 & 415.02 &  4.78 & 3.25 & 13.88 \\
				                    & PSYNTH    & ($N_u$=45)  & 306.90 & 223.51 & 409.92 &  2.34 & 1.54 &  3.95 \\
				\hlineB{2}          
				\multirow{4}{*}{FR} & SRS       & ($N_u$=388) & 301.83 &  99.89 & 612.13 & 18.32 & 0.34 & 104.97 \\
				                    & PSMALL    & ($N_u$=321) & 308.15 & 159.64 & 568.67 & 12.24 & 3.48 &  44.94 \\
				                    & EXTPSYNTH & ($N_u$=321) & 308.38 & 154.07 & 544.34 & 11.34 & 3.60 &  40.91 \\
				                    & PSYNTH    & ($N_u$=403) & 307.82 & 166.01 & 444.29 & 4.65  & 2.56 &  62.51 \\
				\hline \hline
			\end{tabular}
		}%
	\end{center}
\end{table}

% ----------------------------------------------------------------------- %
% ----------------------------------------------------------------------- %
\subsection{Estimation error}
\label{sec:esterr}

On both small area levels, the design-unbiased estimators \psmall{} and \extpsynth{} led to a substantial reduction in the estimation error compared to the SRS estimator (Fig. \ref{fig:disterrors}). On the FA level, the SRS estimator yielded an estimation error of 6.7\% on average compared to 5.2\% and 4.8\% under the \extpsynth{} and \psmall{} respectively (Table \ref{tab:estres}). The cumulative error distribution (Fig. \ref{fig:disterrors}, left) reveals that under the SRS estimator, errors less than 5\% were achieved for 17\% of the FA units (8 of 45). This proportion could be increased to 62\% (28 FA units) and 73\% (33 FA units) by application of the \psmall{} and \extpsynth{} estimator. 95\% of all estimates exhibited errors less than 9.5\% under the SRS estimator and less than 6.6\% when using \psmall{} or \extpsynth{}. Estimation errors higher than 10\% only appeared twice for each of the three estimators.\par
Although the estimation errors were substantially larger overall on the FR level compared to the FA level due to smaller sample sizes, the error reduction from SRS by \psmall{} and \extpsynth{} were even more pronounced (Fig. \ref{fig:disterrors}, right). The average error under the SRS estimator was 18.3\%, while it was 11.3\% and 12.2\% under \psmall{} and \extpsynth{} (Table \ref{tab:estres}). Errors smaller than 10\% were achieved for 15\% of the FR units by the SRS estimator, and for 46\% by the \psmall{} and \psynth{} estimator. 95\% of the 321 FR units where \psmall{} and \extpsynth{} could be applied exhibited errors less than 20\%. In comparion, the SRS estimates resulted in errors less than 36.6\% for 95\% of the 388 FR units.

\begin{figure}[H]
	\centering
	\resizebox{1\hsize}{!}{\includegraphics{fig/error_distr_foa_fu.png}}
	\caption{Cumulative distribution of estimation errors under SRS, \psmall{}, \extpsynth{} and the \psynth{} estimator. \textit{Left}: Results for the 45 FA units. \textit{Right}: Results for the 388 (SRS), 321 (\psmall{} / \extpsynth{}) and 403 (\psynth{}) FR units.}
	\label{fig:disterrors}
\end{figure}

@@

Figure \ref{fig:disterrors} reveals that the error distribution of \psmall{} and \extpsynth{} are very similar, with \psmall{} showing marginally higher estimation errors. In order to investigate the differences between \psmall{} and \extpsynth{}, we compared the g-weight variances of both estimators for all 321 FR units (fig. \ref{fig:compvar}, left). As obvious, \psmall{} yielded slightly larger variances for the vast majority of the estimates. As addressed in section \ref{sec:extpsynth}, one possible explanation for such differences was the effect of one or more cluster not entirely being included in a small area unit, as this would constitute a violation of the \extpsynth{} estimator. This violation was actually observed in 155 of the 321 FR units (48\%).

However, the affected FR units (depicted in red diamonds, fig. \ref{fig:compvar}) did not show a significant divergence from the \psmall{} variances with respect to the remaining unaffected FR units. The variance differences between the two estimators were thus due to the mathematical formulations of the \psmall{} and \extpsynth{} estimator, which are asymptotically equivalent only under large terrestrial sample sizes $n_{2,G}$ within the small area \citep[pp.17--18]{mandallaz2016}. An additional comparison of the absolute differences in the g-weight variance (fig. \ref{fig:compvar}, right) revealed that large divergences did in fact particularly occur for small area units with small terrestrial sample sizes ($n_{2,G} \leq 5$). The differences decreased with increasing sample size and thus confirmed the asymptotic relationship between the two estimators. However, a comparison of the confidence intervals of \psmall{} and \extpsynth{} revealed that the variance differences did not lead to statistically significant point estimates.\par

@@

% ----------------------------------------------------------------------- %
% ----------------------------------------------------------------------- %
\subsection{Variance reduction compared to SRS}
\label{sec:gain_eval}

The variance reduction relative to SRS for \psmall{} and \extpsynth{} are described in Figure \ref{fig:gain} and Table \ref{tab:gain}. A direct comparison of the variances within the small area units revealed that the application of the design-unbiased estimators (\psmall{}, \extpsynth{}) led to a variance reduction compared to SRS in all FA units. In 75\% of the FA units, the \extpsynth{} estimator was able to reduce the variance by up to 54.1\%. The reduction in variance can also be expressed in the relative efficiency values, which were 2.02 on average and ranged between 1.18 and 4.13 on the FA level. On FR level, the reduction in variance and the relative efficiencies reached even higher values (Table \ref{tab:gain} and Fig. \ref{fig:gain}). The \psmall{} estimator again yielded slightly lower variance reductions and relative efficiencies due to the generally smaller variances of the \extpsynth{} estimator (section \ref{sec:comp}).

\begin{figure}[H]
	\centering
	\resizebox{0.73\hsize}{!}{\includegraphics{fig/compare_onephase_extpsynth_psmall.png}}
	\caption{Cumulative distribution of variance reduction by the PSMALL and EXTPSYNTH compared to the SRS estimator for the  45 FA and 321 FR units.}
	\label{fig:gain}
\end{figure}

\begin{table}[H]
	\begin{center}
		\caption{Descriptive summary of variance reduction compared to SRS and relative efficiencies on the two forest district levels. $N_u$: number of evaluated small area units.}
		\vspace{0.2cm}
		\label{tab:gain}
		{\small %
			\begin{tabular}{c|l c|c c c|c c c} %8cols
				\hlineB{1}
				\multirow{2}{*}{District level} & \multicolumn{2}{c|}{\multirow{2}{*}{Estimator}} & \multicolumn{3}{c|}{Variance reduction [\%]} & \multicolumn{3}{c}{relative efficiency} \\
				\cline{4-9} & & & mean & min & max & mean & min & max \\
				\hline \hline
				\multirow{2}{*}{FA} & PSMALL    & ($N_u$=45)  & 33.51 &  2.6 & 72.5 & 1.74 & 1.03 & 3.64 \\
				& EXTPSYNTH & ($N_u$=45)  & 43.30 & 15.7 & 75.8 & 2.03  & 1.18 & 4.13 \\
				\hlineB{2}          
				\multirow{2}{*}{FR} & PSMALL    & ($N_u$=321) & 12.48 & -1203.9 & 96.8 & 2.54 & 0.08 & 31.61  \\
				& EXTPSYNTH & ($N_u$=321) & 24.75 & -892.7  & 97.0 & 2.95 & 0.10 & 33.70 \\
				\hline \hline
			\end{tabular}
		}%
	\end{center}
\end{table}

Cases also occurred on the FR level where one or both two-phase estimators produced larger variance values than under the SRS estimator. This happened in 19\% of the FR units under the \extpsynth{}, and in 24\% of the FR units under the \psmall{} estimator. One possible reason for this was supposed to be a large residual variance due to a poor performance of the regression model within the small area unit. In order to investigate this hypothesis, we analyzed the three variance terms of the \psmall{} estimator (eq. \ref{eq:psmallvar}), i.e. the variance introduced by the uncertainty of the regression coefficients (term 1), the variance caused by estimating the auxiliary means (term 2), and the variance of the model residual (term 3). In general, the residual term is expected to make the largest contribution to the overall variance since it's sample size is based on $n_{2,G}$ whereas the auxiliary term and the coefficient term are based on larger sample sizes, i.e. $n_{1,G}$ and $n_2$ respectively. Figure \ref{fig:fail} illustrates the share of the overall variance by the residual term of the \psmall{} estimator scaled by the overall percentage reduction or increase of the variance compared to SRS for various small area sample sizes $n_{2,G}$.\par

The residual term generally constitutes the dominating part of the \psmall{} variance (around 84\% on average). Although high residual term dominance does not necessary indicate that the \psmall{} variance will be disproportionately large, as apparent from Figure \ref{fig:fail} (\textit{right}), the vast majority of the small areas where the \psmall{} variance was larger than the SRS variance had residual terms contributing over 75\% to the overall \psmall{} variance. Furthermore, the magnitudes of the worst cases tended to occur in lower sample sizes. For example, of the FR units that saw variance increases where $n_{2,G}=4$, the average increase was 272\%, compared to 62\% for FR units with $n_{2,G} > 4$ (Fig. \ref{fig:fail}, \textit{left}).  In comparison, the magnitude of the variance decreases were far more homogeneous than for the variance increases regardless of terrestrial sample size. Since $n_{2,G}$ is the same for \psmall{} and SRS, this implies that the sum of square residuals for the model are likely larger than the sum of square local densities for the clusters in $s_{2,G}$ indicating the presence of outliers with large residuals in the problematic small areas. This situation is likely to arise when there was forest loss after the ALS scanning but before the terrestrial survey year. % [check that]

%As expected, the residual term generally constitutes the dominating part of the \psmall{} variance (around 84\% on average). However, a high proportion of the residual variance term seems not to be the driver for large \psmall{} variances, as apparent from Fig. \ref{fig:fail} (\textit{right}). The FR units where the \psmall{} estimator produced larger variances than the SRS estimator did not systematically differ from the majority of the FR units where \psmall{} performed better than SRS. However, FR units with exceptionally large variance increases compared to SRS particularly occurred under small terrestrial sample sizes of $n_{2,G} = 4$ (fig. \ref{fig:fail}, \textit{left}). These FR units exhibited a 272\% average increase of the SRS variance, compared to 62\% for the critical FR units with $n_{2,G} > 4$. In comparison, the achievable reduction in SRS variances compared to SRS were not considerably impacted by the terrestrial sample size (fig. \ref{fig:fail}, \textit{right}). The average reduction of these units was around 50\% under sampling sizes of both $n_{2,G} = 4$ and $n_{2,G} > 4$.
